---
title: "Improving a UK Opinion Chart"
author: "Derwin McGeary"
date: "10/07/2015"
output: html_document
---

# Introduction
## Motivation
A survey was done by YouGov asking the British public which groups in society they considered the two main political parties to be closer to.
The graphs in [the Independent](http://i100.independent.co.uk/article/the-dna-of-british-politics-in-one-chart--ZkrJXf94wx) and on [YouGov (first chart)](https://yougov.co.uk/news/2015/10/04/party-working-people/) have some serious shortcomings.  Specifically, the Independent uses one side of the horizontal chart for Labour, and one for the Conservatives, and uses colour to code for "Close" and "Not Close" (both percentage variables in the same category). The YouGov chart is much more legible, but it counterintuitively displays bars **further away** from the group names to indicate **proximity**.

# Code
## Downloading the data
We don't want to download multiple times when testing, we **do** want people to be able to run this very document so a convenience function to check, download and read in the data. This does require that `pdftotext` is installed.
```{r}
loadfile <- function(myurl) {
  library(downloader)
  if(!file.exists("Report.pdf")) {
    download(myurl, "Report.pdf") }
  if(!file.exists("Report.txt")) {
    system("pdftotext -layout Report.pdf") }
  readLines("Report.txt")
}
```
## Scraping the data
The results from the above function are rather messy (just a text document in horizontal strips).  All the tables are roughly similar, so this function takes a line number for the start of the table and extracts the relevant information
```{r}
raw2stats <- function(rawdata, party, start) {
# conservative is 2, labour is 1... let's not get fancy
  thisparty <- ifelse(party==2,"Conservative","Labour")
  # Category name, CLOSE, NOT CLOSE
  coords <- as.data.frame(rbind(c(start,1,24), c(start+3,81,83), c(start+6,81,83)))
  colnames(coords) <- c("line","start","end")
  group <- substr(rawdata[start],1,24)
  liq <- suppressWarnings(as.numeric(unlist(strsplit(rawdata[sum(start,3)]," ")), warnings=FALSE))
  lir <- suppressWarnings(as.numeric(unlist(strsplit(rawdata[sum(start,6)]," ")), warnings=FALSE))
  liq <- liq[!is.na(liq)]
  lir <- lir[!is.na(lir)]
  close <- liq[2]
  notclose <- lir[2]
  #framedata <- as.data.frame(rbind(with(coords, substr(rawdata[line],start,end))))
  framedata <- cbind(thisparty, group, close, notclose)
  colnames(framedata) <- c("Party", "Group", "Close", "NotClose")
  framedata
}
```
## Plotting
The plotting function is designed to produce a horizontal bar chart where physical proximity to the label indicates perceived closeness to the group. Red and blue code for Labour and Conservative respectively.
```{r}
plotme <- function(dataset) {
  library(ggplot2)
  library(xkcd)
  p <- ggplot(data=dataset, aes(x=Group, y=Number, fill=Party)) + 
    geom_bar(stat="identity", position=position_dodge()) +
    scale_y_continuous(breaks=seq(-90,90,10),label=abs(seq(-90,90,10))) +
    scale_fill_manual(values=c("#9999FF","#FF5555")) +
    scale_colour_manual(values=c("#555555","#FF5555")) +
    ggtitle(expression(atop("How close is each party to the following groups?", atop("      % Consider Close <--------------------       -------------------> % Consider Not Close","")))) +
    ylab("") +
    xlab("") +
    theme_bw() +
    geom_hline(aes(yintercept=0),colour="#FFFFFF") +
    coord_flip()
  print(p)
}
```

The rest of the program does the bookkeeping to select the subtables and reshape and sort the dataframe before plotting.
```{r}
# Read the data in as a character vector of lines
reporturl <- "https://d25d2506sfb94s.cloudfront.net/cumulus_uploads/document/dmi10ux8ie/InternalResults_151002_PartyGroups_Website.pdf"
raw <- loadfile(reporturl) 

# Convert to a Dataframe
conservative <- c(21,29,37,45,72,80,88,96,104,131,139,147) # Line numbers for 
labour <- c(179,187,195,203,230,238,247,255,263,289,297,305) # each sub table in the pdf
starts <- as.data.frame(cbind(conservative,labour))
output <- data.frame()
for(i in starts[,1]) output <- rbind(output,raw2stats(raw,2,i))
for(i in starts[,2]) output <- rbind(output,raw2stats(raw,1,i))

# Tidy it up a bit
library(tidyr)
output$Close <- (-1)*as.numeric(as.character(output$Close)) # We make the "Close" value negative...
output$NotClose <- as.numeric(as.character(output$NotClose)) # ...  so that it's physically closer to the label
output <- gather(output, Measure, Number, Close:NotClose)
output$Group <- factor(output$Group, levels=unique(output$Group[order(output$Party,output$Measure, (-1)*output$Number)]))

# ... and plot!
plotme(output)
```